### 2021-11-08 article/article.tex
Article update 8

\section{Title}
Spinsim: a GPU optimized simulator of spin-half and spin-one quantum systems

@@

\section{Paper Author Roles and Affiliations}
1. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Primary author of the released packages.\\
2. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: Faculty of Physics, University of Vienna, 1010 Vienna, Austria.\\
	Author of first version of code.\\
3. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: School of Physics, University of Sydney, NSW 2006, Australia.\\
	Optimization and extension to spin-one of first version of code.\\
4. School of Molecular Sciences, La Trobe University, PO box 199, Bendigo, Victoria 3552, Australia.\\
	Present address: Q-CTRL Pty. Ltd.\\
	Original conception of first version of code.\\
5. Department of Electrical and Computer Systems Engineering, Monash University, Victoria 3800, Australia.\\
	Advice on numerical analysis.\\
6. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Original conception of released version of algorithm.

\section{Abstract}
	The \emph{Spinsim} \emph{python} package simulates spin-half and spin-one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda}, which is an \emph{LLVM} (Low Level Virtual Machine) compiler for \emph{Nvidia Cuda} compatible systems using GPU parallelisation. 
	Along with other optimisations, this allows for speed improvements of up to four orders of magnitude while keeping staying just as accurate.
	\emph{Spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use-case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin-half or spin-one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin-one systems of nitrogen vacancy centres (NVCs) and BECs.

\section{Keywords}
Time dependent Schr\"odinger equation; Quantum; Physics; Spin one; Spin half; Integrator; Exponentiator; GPU; Parallel computing; Solver; python; numba;

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing\cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin-half (ie spin quantum number of $ \frac12 $, also referred to as a qubit), is quantized into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin-one quantum system (ie spin quantum number of $ 1 $, also referred to as a quitrit), which is quantized into three quantum spin levels.

	The design of sensing protocols requires many steps of verification, including simulation.
	This is especially important, since running real experiments can be expensive and time consuming, and thus it is more practical to debug such protocols quickly and cheaply on a computer.
	In general, any design of experiments using spin systems could benefit from a fast, accurate method of simulation.

	% \texttt{AtomicPy}\cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package is based, and standard differential equation solvers (such as \emph{Mathematica}'s function \texttt{NDSolve}) to solve the Schr\"odinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin-half for a faster execution time, at the cost of modelling all effects of the system.
	However, these methods are not completely optimized for our use-case, and therefore come with some issues.

	First, while the execution time for these solvers is acceptable for running a small number of experiments, for certain experiments involving large arrays of independent atom clouds (which require many thousands of simulations to be run), this time accumulates to the order of many hours, or even multiple days.

	Second, the Schr\"odinger equation has the geometric property of being norm persevering.
	In other words, the time-evolution operator for a system between two points in time must be unitary.
	As such, numerical solutions to the Schr\"odinger equation should also preserve this property.
	For many numerical methods like those in the Runge Kutta family, the approximations used might not be norm preserving, and the evaluated quantum state may diverge towards an infinite norm, or converge to zero if run for many iterations.

	Third, our system (and similar spin systems) can be very oscillatory.
	In standard conditions for our application, the expected spin projection of a system that we want to solve for can rotate in physical space (alternatively viewed as a point rotating around an abstract object known as a \emph{Bloch sphere}) at a rate of 700kHz.
	Standard integration methods require very small time-steps in order to accurately depict these oscillations.

	Given the recent boom in machine learning, many research computers are now equipped with advanced graphics processing units (GPUs).
	Their many cores, which can range from hundreds to tens of thousands in number, allows them to run some highly parallel algorithms much faster than a central processing unit (CPU).
	Examples include calculating colors for many pixels on a screen in the titular graphics processing, or the weights in a large neural network.
	By finding ways to parallelize the problem of solving quantum spin systems, we can use the many cores of a GPU to our advantage in this context as well.

@@

		Rather than parametrizing the problem in terms of the matrix elements of $ H(t) $, we consider time varying real coefficients in a linear combination of fixed operators,
		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} \omega_j(t) \mathcal{A}_j.
		\end{align}
		Here we exclude the identity so that the Hamiltonian is traceless, which corresponds to choosing a physically meaningless energy zero point.
		It is well known that a charged spin-half system with magnetic moment $ \overrightarrow{\mu} $, and gyromagnetic ratio $ \gamma $, in a magnetic field $ \overrightarrow{B}(t) $, has Hamiltonian
		\begin{align}
			H(t) &= -\overrightarrow{B}(t)\cdot \overrightarrow{\mu}\\
			% &= -\gamma \overrightarrow{B}(t)\cdot \overrightarrow{J}\\
			&= -\gamma \left(B_x(t) J_x + B_y(t) J_y + B_z(t) J_z\right)\\
			% &= \left(-\gamma B_x(t) J_x\right) + \left(-\gamma B_y(t) J_y\right) + \left(-\gamma B_z(t) J_z\right)\\
			&= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z.
		\end{align}
		Here $ J_x $, $ J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices\cite[(p169)]{j_j_sakurai_jun_john_modern_2011} halved.
		% \supervisordisagreement{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices halved.
		% }{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators
		% 	\begin{align}
		% 		J_x &= \frac12\begin{pmatrix}
		% 			0 & 1 \\
		% 			1 & 0
		% 		\end{pmatrix},
		% 		&J_y &= \frac12\begin{pmatrix}
		% 			0 & -i \\
		% 			i &  0
		% 		\end{pmatrix},
		% 		&\textrm{and }J_z &= \frac12\begin{pmatrix}
		% 			1 &  0 \\
		% 			0 & -1
		% 		\end{pmatrix},\label{eq:spin_half_operators}
		% 	\end{align}
		% 	equal to the Pauli matrices halved.
		% }
		Because of the equivalence of the magnetic field components $ B_j(t) $ and the $ \omega_j(t) $, we henceforth refer to the latter as field functions.

		In the spin-one case there is no standard basis of operators $ \mathcal{A}_j $.
		Choices include the Gell-Mann matrices\cite{gell-mann_symmetries_1962}, and multiple dipole-quadrupole bases\cite{hamley_spin-nematic_2012, di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra $ \mathfrak{su}(3) $, which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group $ SU(3) $) in the spin-one system.
		We focus on a particular subfamily of of spin-one systems for which the Hamiltonian is a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x $, $ J_y $ and $ J_z $, (labelled $ L_x $, $ L_y $ and $ L_z $ in Reference~\cite{hamley_spin-nematic_2012}) and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,% \todo{Describe this subspace better}

@@

		Note that $ Q $ is proportional to $ Q_{zz} $\cite{hamley_spin-nematic_2012} and $ Q_0 $\cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		The $ J_x $, $ J_y $, $ J_z $ and $ Q $ are the only operators necessary to simulate many spin-one quantum systems in arbitrary bias fields, but with single-photon coupling.
		This includes quadratic Zeeman splitting described by the Breit-Rabi formula\cite{mockler_atomic_1961} important to experiments in our lab.
		The \emph{Spinsim} simulator can also be configured to solve a general spin-one system by setting the Hamiltonian to an arbitrary point in $ \mathfrak{su}(3) $ using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) =& \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{u1}(t) U_1 + \omega_{u2}(t) U_2 + \omega_{v1}(t) V_1 + \omega_{v2}(t) V_2.
		\end{align}
		Here the additional operators are those defined in\cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the $ U_1, U_2, V_1 $ and $ V_2 $ operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% The matrices $ J_x, J_y, J_z $ are regular spin operators, and $ Q $ is a quadrupole operator. Note that $ Q $ is proportional to $ Q_{zz} $ as defined by Hamley et al\cite{hamley_spin-nematic_2012}, and $ Q_0 $ as defined by Di et al\cite{di_dipolequadrupole_2010}.
		% INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin-one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		As well as integrating the Schr\"odinger equation, \emph{Spinsim} also has the functionality to calculate the expected spin projection $ \langle \overrightarrow{J}\rangle (t) $ of a system from its state.
		In an experimental setting, one cannot measure the value of the state itself, and must instead measure observables such as spin projection.
		If this is done for an ensemble of systems, the average spin projection over all systems will converge to the expected value as calculated here.
		As an alternative to solving the Schr\"odinger equation, if one is only interested in the dynamics of the expected spin projection, one can instead solve for $ \langle \overrightarrow{J}\rangle (t) $ directly using the Bloch equations\cite{bloch_nuclear_1946}.
		These are three coupled real equations as compared to three coupled complex equations for the spin-one Schr\"odinger equation.
		However, the Bloch equations cannot be used to compute $ \psi(t) $, and can only fully model spin-half systems (and spin-one systems approximated to spin-half).
		Thus, with \emph{Spinsim} we choose to solve the full Schr\"odinger equation instead.

@@

	\subsection{Unitary time evolution and the Magnus expansion}
		Since $H(t)$ is Hermitian, it follows from Equation~\eqref{eq:schroedinger} that $\|\psi(t)\|^2 = \|\psi(t_0)\|^2$ for all $t$.
		Therefore it is possible to write $ \psi(t) $ in terms of a unitary transformation $ \mathcal{U}(t, t_0) $ of the state $ \psi(t_0) $, for any time $ t_0 $, that is, $ \psi(t) = \mathcal{U}(t, t_0)\psi(t_0) $.
		It then follows from Equation~\eqref{eq:schroedinger} that $ \mathcal{U}(t, t_0) $ also follows a Schr\"{o}dinger equation,
		\begin{align}
			i\hbar\frac{\mathrm{d}\mathcal{U}(t, t_0)}{\mathrm{d}t} &= H(t)\mathcal{U}(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation~\eqref{eq:schroedinger} for $ \psi(t) $ given a particular $ \psi(t_0) $, one only needs to solve Equation~\eqref{eq:unitary_schroedinger} for $ \mathcal{U}(t, t_0) $.
		% Note that because $ \mathcal{U}(t, t_0) $ is unitary, the calculated value of $ \psi(t) $ using this method will not change in magnitude, conserving probability.
		Since $ \mathcal{U}(t, t_0) $ is unitary, it can be written as a matrix exponential of a anti-Hermitian (also termed skew-Hermitian) matrix.
		If the Hamiltonian is a of constant value $ H(t) = \mathcal{H} $, then this exponential is % \note{Should we use H, ie no argument, or H(t'), ie arbitrary argument, to represent constancy?} % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i (t - t_0) \mathcal{H}).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series\cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for $ \mathcal{U}(t, t_0) $ is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The well known Dyson series\cite{kalev_integral-free_2020} gives an explicit expression for $ \mathcal{U}(t, t_0) $ in terms of multiple time integrals over nested time commutators of $ H(t') $.
		While the Dyson series has recently been used numerically\cite{kalev_integral-free_2020}, it is challenging to work with because once truncated, it is in general no longer unitary.

>>>>

\section{Title}
% Spinsim: a GPU optimized simulator of spin-half and spin-one quantum systems
Spinsim: a GPU optimized python package for simulating spin-half and spin-one quantum systems

@@

\section{Paper Author Roles and Affiliations}
1. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Primary author of the released packages.\\
2. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: Faculty of Physics, University of Vienna, 1010 Vienna, Austria.\\
	Author of first version of code.\\
3. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: School of Physics, University of Sydney, NSW 2006, Australia.\\
	Optimization and extension to spin-one of first version of code.\\
4. School of Molecular Sciences, La Trobe University, PO box 199, Bendigo, Victoria 3552, Australia.\\
	Present address: Q-CTRL Pty. Ltd.\\
	Original conception of first version, and advice on current version of code.\\
5. Department of Electrical and Computer Systems Engineering, Monash University, Victoria 3800, Australia.\\
	Advice on numerical analysis.\\
6. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Original conception of released version of algorithm.

\section{Abstract}
	The \emph{Spinsim} \emph{python} package simulates spin-half and spin-one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda}, which is an \emph{LLVM} (Low Level Virtual Machine) compiler for \emph{Nvidia Cuda} compatible systems using GPU parallelization. 
	Along with other optimizations, this allows for speed improvements of up to four orders of magnitude while staying just as accurate, compared to industry standard packages.
	It is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use-case for the \emph{Spinsim} will be to simulate quantum sensing-based ultracold atom experiments for the Monash University School of Physics \& Astronomy spinor Bose-Einstein condensate (spinor BEC) lab, but we anticipate it will be useful in simulating any range of spin-half or spin-one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin-one systems of nitrogen vacancy centres (NVCs), ultracold atoms, and BECs.

\section{Keywords}
Time dependent Schr\"odinger equation; Quantum; Physics; Spin one; Spin half; Integrator; Exponentiator; GPU; Parallel computing; Solver; Python; Numba; Sensing;

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing~\cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin-half (ie spin quantum number of $ \frac12 $, also referred to as a qubit), is quantized into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin-one quantum system (ie spin quantum number of $ 1 $, also referred to as a quitrit), which is quantized into three quantum spin levels.

	The design of sensing protocols requires many steps of verification, including simulation.
	This is especially important, since running real experiments can be expensive and time consuming, so it is more practical to debug such protocols quickly and cheaply on a computer.
	In general, any design of experiments using spin systems could benefit from a fast, accurate method of simulation.

	% \texttt{AtomicPy}~\cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package, \emph{Spinsim}, is based, and standard differential equation solvers (such as \emph{Mathematica}'s~\cite{wolfram_research_inc_mathematica_2020} function \texttt{NDSolve}) to solve the Schr\"odinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin-half for a faster execution time, at the cost of not modelling all effects in the system.
	However, these methods are not completely optimized for our use-case, and therefore come with some issues.

	First, while the execution time for these solvers is acceptable for running a small number of experiments, for certain experiments involving large arrays of independent atom clouds (which require many thousands of simulations to be run), this time accumulates to the order of many hours, or even multiple days.
	Given the recent boom in machine learning, many research computers are now equipped with modern graphics processing units (GPUs).
	Their many cores, which can range from hundreds to tens of thousands in number, allows them to run some highly parallel algorithms much faster than a central processing unit (CPU).
	Examples include calculating color values for many pixels on a screen in the titular graphics processing, or the weights in a large neural network.
	While there do exists parallelized algorithms to solve general differential equations~\cite{lions_resolution_2001}, there are properties specific to the spin system that allow for much more fast and elegant parallelization in our use-case.
	We use these properties (see \emph{\nameref{sec:parallelization}}) to write a parallel, GPU based solver that runs faster than industry standard non-parallel algorithms by three orders of magnitude.

	Second, the Schr\"odinger equation has the geometric property of being norm persevering.
	In other words, the time-evolution operator for a system between two points in time must be unitary.
	As such, numerical solutions to the Schr\"odinger equation should also preserve norms.
	For many numerical methods like those in the Runge Kutta family, the approximations used are in general not norm preserving, so the evaluated quantum state may diverge towards an infinite norm, or converge to zero if run for many iterations.
	To avoid this, we use a geometric integration technique for our solver (see \emph{\nameref{sec:magnus}}).

	Third, our system (and similar spin systems) can be very oscillatory.
	In standard conditions for our application, the expected spin projection of a system that we want to solve for can rotate in physical space (alternatively viewed as a point rotating around an abstract object known as a \emph{Bloch sphere}) at a rate of 700kHz.
	Standard integration methods require very small time-steps in order to accurately depict these oscillations.
	For this reason we use a fourth order Magnus expansion~\cite{magnus_exponential_1954,blanes_magnus_2009} based solver for our geometric integrator to make larger time-steps accurate (see \emph{\nameref{sec:magnus}}).
	We also use a dynamically changing rotating frame to slow down the effective rotations in the system, and thus reduce the size of integration steps being done, making each step more accurate (see \emph{\nameref{sec:rotating_frame}}).

	% By finding ways to parallelize the problem of solving quantum spin systems, we can use the many cores of a GPU to our advantage in this context as well.

	\subsection{Recent literature}
		While we benchmark \emph{Spinsim} only against industry standard integrators, we also note that application specific quantum numerical simulator packages are regularly being released.
		However, none of these packages are designed specific to quantum sensing, and so these releases are not suitable for simulations highly accurate at high time resolutions required in that field.

		In quantum optimal control (QOC), researchers look at how to design robust pulse sequences to control spin systems using machine learning techniques.
		Particularly, this means that simulators used here require automatic differentiation~\cite{griewank_who_2012}, which \emph{Spinsim} does not currently support.
		A main use of QOC is to design control sequences in quantum computers.
		Many of these pulse sequences are trains of hard pulses (which are physically constant bursts of radiation), which, if using the accuracy reducing rotating wave approximation (RWA), can have simplified time evolution operators.
		This paradigm, used in the recent \emph{Qulacs}~\cite{suzuki_qulacs_2021}, \emph{QOpt}~\cite{teske_qopt_2021} and \emph{QSim}~\cite{isakov_simulations_2021}, is not suitable for accurately simulating more complicated (especially when looking from the lab frame of reference) pulse shapes found in quantum sensing.
		However, all of these packages do allow for simulating entanglement (important for quantum computing, but not for all sensing applications), which \emph{Spinsim} does not and, like \emph{Spinsim}, all have versions optimised for parallelisation (although \emph{QSim} does sacrifice accuracy here for the sake of speed again, by using single-precision floats).
		Dalgaard and Motzoi~\cite{dalgaard_fast_2021} have recently proposed Magnus integration techniques for use in QOC, but have not provided any open source implementation to do so.

		% Three qubit such packages were published recently, being \emph{PULSEE}~\cite{candoli_pulsee_2021}, \emph{QOpt}~\cite{teske_qopt_2021} and ~\cite{isakov_simulations_2021}.
		% While primarily focusing on being a complete toolkit for NMR simulations and quantum control optimization respectively, both contain Schr\"odinger equation solvers.
		% Furthermore, neither package can take advantage of GPU parallelization.

		Another application of simulations of spin systems is nuclear magnetic resonance (NMR).
		In this field, the simulator part of the \emph{PULSEE}~\cite{candoli_pulsee_2021} package uses up to the 3rd order Magnus expansion just once to solve the complete system, rather than using it many times as a time stepping method.
		Nested integrals of commutators are numerically evaluated on a CPU using Euler integration.
		This is likely to be slow (due to the Euler-integrated nested integrals of commutators) and inaccurate (due to using the expansion over a large time step), and potentially not even convergent in general (also due to the large time step); although it may suffice for the kinds of problems this NMR simulation package is looking at.

		% Designed for optimising quantum control protocols, the \emph{QOpt} package works in the paradigm of a time series of instructions, rather than the analytic forms used in \emph{Spinsim}.
		% Given the solid grid of time series, this package cannot take advantage of the commutator free Magnus based techniques that \emph{Spinsim} employs.
		% It may be able to model hard pulse sequences accurately (prevalent in quantum control), but will likely not be as accurate when modelling a continuous magnetic environment of a spin system.

		% Finally, also just recently, Dalgaard and Motzoi~\cite{dalgaard_fast_2021} propose the use of Magnus expansion based techniques for simulations in quantum control optimization.
		% However, they do not provide an implementation of their work.

@@

		Rather than parametrizing the problem in terms of the matrix elements of $ H(t) $, we consider time varying real coefficients in a linear combination of fixed operators,
		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} \omega_j(t) \mathcal{A}_j.
		\end{align}
		Here we exclude the identity so that the Hamiltonian is traceless, which corresponds to choosing a physically meaningless energy zero point.
		It is well known that a charged spin-half system with magnetic moment $ \overrightarrow{\mu} $, and gyromagnetic ratio $ \gamma $, in a magnetic field $ \overrightarrow{B}(t) $, has Hamiltonian
		\begin{align}
			H(t) &= -\overrightarrow{B}(t)\cdot \overrightarrow{\mu}\\
			% &= -\gamma \overrightarrow{B}(t)\cdot \overrightarrow{J}\\
			&= -\gamma \left(B_x(t) J_x + B_y(t) J_y + B_z(t) J_z\right)\\
			% &= \left(-\gamma B_x(t) J_x\right) + \left(-\gamma B_y(t) J_y\right) + \left(-\gamma B_z(t) J_z\right)\\
			&= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z.
		\end{align}
		Here $ J_x $, $ J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices~\cite[(p169)]{j_j_sakurai_jun_john_modern_2011} halved.
		% \supervisordisagreement{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices halved.
		% }{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators
		% 	\begin{align}
		% 		J_x &= \frac12\begin{pmatrix}
		% 			0 & 1 \\
		% 			1 & 0
		% 		\end{pmatrix},
		% 		&J_y &= \frac12\begin{pmatrix}
		% 			0 & -i \\
		% 			i &  0
		% 		\end{pmatrix},
		% 		&\textrm{and }J_z &= \frac12\begin{pmatrix}
		% 			1 &  0 \\
		% 			0 & -1
		% 		\end{pmatrix},\label{eq:spin_half_operators}
		% 	\end{align}
		% 	equal to the Pauli matrices halved.
		% }
		Because of the equivalence of the magnetic field components $ B_j(t) $ and the $ \omega_j(t) $, we henceforth refer to the latter as field functions.

		In the spin-one case there is no standard basis of operators $ \mathcal{A}_j $.
		Choices include the Gell-Mann matrices~\cite{gell-mann_symmetries_1962}, and multiple dipole-quadrupole bases~\cite{hamley_spin-nematic_2012, di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra $ \mathfrak{su}(3) $, which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group $ SU(3) $) in the spin-one system.
		We focus on a particular subfamily of of spin-one systems for which the Hamiltonian is a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x $, $ J_y $ and $ J_z $, (labelled $ L_x $, $ L_y $ and $ L_z $ in Reference~\cite{hamley_spin-nematic_2012}) and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,% \todo{Describe this subspace better}

@@

		Note that $ Q $ is proportional to $ J_z^2 $ (up to a change in energy zero point), and $ Q_{zz} $~\cite{hamley_spin-nematic_2012} and $ Q_0 $~\cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		The $ J_x $, $ J_y $, $ J_z $ and $ Q $ are the only operators necessary to simulate many spin-one quantum systems in arbitrary bias fields, but with single-photon coupling.
		This includes quadratic Zeeman splitting described by the Breit-Rabi formula~\cite{mockler_atomic_1961} important to experiments in our lab.
		The \emph{Spinsim} simulator can also be configured to solve a general spin-one system by setting the Hamiltonian to an arbitrary point in $ \mathfrak{su}(3) $ using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) =& \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{u1}(t) U_1 + \omega_{u2}(t) U_2 + \omega_{v1}(t) V_1 + \omega_{v2}(t) V_2.
		\end{align}
		Here the additional operators are those defined in~\cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the $ U_1, U_2, V_1 $ and $ V_2 $ operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% The matrices $ J_x, J_y, J_z $ are regular spin operators, and $ Q $ is a quadrupole operator. Note that $ Q $ is proportional to $ Q_{zz} $ as defined by Hamley et al~\cite{hamley_spin-nematic_2012}, and $ Q_0 $ as defined by Di et al~\cite{di_dipolequadrupole_2010}.
		% INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin-one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		As well as integrating the Schr\"odinger equation, \emph{Spinsim} also has the functionality to calculate the expected spin projection $ \langle \overrightarrow{J}\rangle (t) $ of a system from its state.
		In an experimental setting, one cannot measure the value of the state itself, and must instead measure observables such as spin projection.
		If this is done for an ensemble of systems, the average spin projection over all systems will converge to the expected value as calculated here.
		As an alternative to solving the Schr\"odinger equation, if one is only interested in the dynamics of the expected spin projection, one can instead solve for $ \langle \overrightarrow{J}\rangle (t) $ directly using the Bloch equations~\cite{bloch_nuclear_1946}.
		These are three coupled real equations as compared to three coupled complex equations for the spin-one Schr\"odinger equation.
		However, the Bloch equations cannot be used to compute $ \psi(t) $, and can only fully model spin-half systems (and spin-one systems approximated to spin-half).
		Thus, with \emph{Spinsim} we choose to solve the full Schr\"odinger equation instead.

@@

	\subsection{Unitary time evolution and the Magnus expansion}\label{sec:magnus}
		Since $H(t)$ is Hermitian, it follows from Equation~\eqref{eq:schroedinger} that $\|\psi(t)\|^2 = \|\psi(t_0)\|^2$ for all $t$.
		Therefore it is possible to write $ \psi(t) $ in terms of a unitary transformation $ \mathcal{U}(t, t_0) $ of the state $ \psi(t_0) $, for any time $ t_0 $, that is, $ \psi(t) = \mathcal{U}(t, t_0)\psi(t_0) $.
		It then follows from Equation~\eqref{eq:schroedinger} that $ \mathcal{U}(t, t_0) $ also follows a Schr\"{o}dinger equation,
		\begin{align}
			i\hbar\frac{\mathrm{d}\mathcal{U}(t, t_0)}{\mathrm{d}t} &= H(t)\mathcal{U}(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation~\eqref{eq:schroedinger} for $ \psi(t) $ given a particular $ \psi(t_0) $, one only needs to solve Equation~\eqref{eq:unitary_schroedinger} for $ \mathcal{U}(t, t_0) $.
		% Note that because $ \mathcal{U}(t, t_0) $ is unitary, the calculated value of $ \psi(t) $ using this method will not change in magnitude, conserving probability.
		Since $ \mathcal{U}(t, t_0) $ is unitary, it can be written as a matrix exponential of a anti-Hermitian (also termed skew-Hermitian) matrix.
		If the Hamiltonian is a of constant value $ H(t) = \mathcal{H} $, then this exponential is % \note{Should we use H, ie no argument, or H(t'), ie arbitrary argument, to represent constancy?} % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i (t - t_0) \mathcal{H}).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series~\cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for $ \mathcal{U}(t, t_0) $ is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The well known Dyson series~\cite{kalev_integral-free_2020} gives an explicit expression for $ \mathcal{U}(t, t_0) $ in terms of multiple time integrals over nested time commutators of $ H(t') $.
		While the Dyson series has recently been used numerically~\cite{kalev_integral-free_2020}, it is challenging to work with because once truncated, it is in general no longer unitary.

		The Magnus series, in contrast, gives $ \mathcal{U}(t, t_0) $ in terms of an exponential of a series of anti-Hermitian operators, specifically
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp\left(\Omega(t, t_0)\right) = \exp\left(\sum_{m = 1}^\infty \Omega_m(t, t_0)\right).
		\end{align}
		As such, the Magnus series explicitly preserves unitarity when truncated~\cite{magnus_exponential_1954}.
		The terms in the Magnus series are integrals over nested time commutators of the Hermitian matrix $ A(t') $, where, in the case of quantum mechanics, $ A(t') = -iH(t') $.
		For instance, the first three terms are~\cite{blanes_magnus_2009}
		\begin{align}
			\Omega_1(t, t_0) &= \int_0^t\mathrm{d}t_1A(t_1)\\
			\Omega_2(t, t_0) &= \frac12\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2[A(t_1), A(t_2)]\\
			\Omega_3(t, t_0) &= \frac16\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2\int_{t_0}^{t_2}\mathrm{d}t_3\left([A(t_1), [A(t_2), A(t_3)]] + [A(t_3), [A(t_2), A(t_1)]]\right),
		\end{align}
		where $ [X, Y] = XY - YX $ is the commutator.
		Note that, when truncated to first order, the Magnus expansion $ \mathcal{U}(t, t_0) = \exp\left(\int_{t_0}^t\mathrm{d}t_1\left(-iH(t_1)\right)\right) $ reduces to Equation~\eqref{eq:exp_sol_of_constant}, with $ H(t) $ approximated by its average value over the interval $ [t_0, t] $.
